Bertrand Russell considers Timaeus to be historically important, but philosophically unimportant. As he wrote,

\begin{quotation}
    ``PLATO'S cosmogony is set forth in the Timaeus, which was translated into Latin by Cicero, and was, in consequence, the only one of the dialogues that was known in the West in the Middle Ages. Both then, and earlier in Neoplatonism, it had more influence than anything else in Plato, which is curious, as it certainly contains more that is simply silly than is to be found in his other writings. As philosophy, it is unimportant, but historically it was so influential that it must be considered in some detail.''
\end{quotation}

In a similar manner, I consider Timaeus to be philosophically dated.

For instance, Bertrand Russell remarked that, “[the] theory of the regular solids, which is set forth in the thirteenth book of Euclid, was, in Plato's day, a recent discovery”. Which highlights a personal pain point I have with regards to Timaeus, and this is, there is very little mathematics that Plato could draw upon, so much so, that Timaeus feels rather impoverished in this regard, what substance there is, feels rather contrived, as I will explain shortly. I’d certainly love to feel something with regards to the mathematics in Timaeus, but as I have said, there is very little that Plato could draw upon, and therefore, in reading Timaeus, your thirst will not be quenched, despite everyone saying as much. 

For instance, with modern mathematics, you can look at a graph of a function in $\mathbb{R}^3$ and see that sometimes you have to regress, to find a more global maxima. It’s so evidently manifested, no-one says it. But it’s interesting because regressing to find a ‘higher’ state, so to speak, is something evolution is incapable of selecting for, it’s something uniquely intrinsic to intelligent life. Maybe you can extrapolate such to life in a philosophical manner; sometimes you have to abandon something, and regress in comforts, or what have you, to find a greater, higher state in life, or as we say in calculus, a global maxima. This meaning is something plainly manifested in many nonlinear functions, and this is just elementary calculus, its just getting your feet wet. 
